---
set_title: !r test_title
output: 
  html_document:
    toc: yes
    toc_float:
      collapsed: true
    toc_depth: 3
    number_sections: true
    self_contained: yes
    css: corp-styles.css
    highlight: pygments
---

```{r, echo=FALSE}
htmltools::img(src = "logo.png",
               alt = 'logo',
               style = 'position:absolute; top:0; right:0; down:20; padding:5px; width:100px; height:auto;')
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<!-- Libraries  -->

```{r libraries, echo=FALSE, warning=FALSE, include=FALSE, message = FALSE}

#load libraries
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(ggplot2)
library(patchwork)
library(rRACES)
library(dplyr)
library(tidyverse)
source("/orfeo/cephfs/scratch/cdslab/ggandolfi/prj_races/REPO_UPDATED/rRACES-examples/report/plotting/utils.R")
source("/orfeo/cephfs/scratch/cdslab/ggandolfi/prj_races/REPO_UPDATED/rRACES-examples/report/plotting/plot_genome_wide.R")
metadata <- read.table(file = "SCOUT_metadata.csv",header = T,sep = "\t")
img_dir <- "SCOUT_images/"
```

```{r params, include = FALSE}
#Load params from YAML if running outside R Markdown
if (!exists("params")) {
  params <- yaml::read_yaml("/orfeo/cephfs/scratch/cdslab/ggandolfi/prj_races/REPO_UPDATED/rRACES-examples/SCOUT/SPN01/params.yml")
}
```

<!-- Read params files -->

```{r assign_variables, include = FALSE, echo =TRUE}
coverage <- params$sequencing$coverage
purity <- params$sequencing$purity
seq_dir <- params$files$seq_res
seq_res_tumour_rds <- paste0(seq_dir,"purity_",purity,"/seq_results_muts_merged_coverage_",coverage,"x.rds")
seq_res_normal_rds <- paste0(seq_dir,"seq_results_muts_merged_coverage_30x.rds")
files_cna <- list.files(params$cna_dir, full.names = T)
sample_names <- sapply(files_cna, function(path) {
  base_name <- basename(path)
  sub("_cna.rds$", "", base_name)
})
number_of_samples <- length(sample_names)
```

<!-- Load data -->

```{r load_data, include = FALSE, echo =TRUE}
phylo_forest <- load_phylogenetic_forest(params$files$phylo_forest)
seq_res_tumour <- readRDS(seq_res_tumour_rds)
seq_res_normal <- readRDS(seq_res_normal_rds)
cna_data <- lapply(list.files(params$cna_dir,recursive = T,full.names = T),readRDS)
```

# General sequencing information

Tumor samples have been sequenced with a coverage of `r params$sequencing$coverage`x and a simulated purity of `r params$sequencing$purity`. The normal sample is simulated from a real individual at a sequencing depth of 30x.

```{r germline_subject}
library(reactable)
library(sparkline)

germ <- phylo_forest$get_germline_subject()
reactable(germ,filterable = FALSE, resizable = T,highlight = T,defaultPageSize = 5,
          bordered = TRUE, striped = TRUE)
```

## Normal sample

Normal sample was sequenced with a coverage of 30X. The plot shows the depth of sequencing of simulated mutations in the normal sample. The \textcolor{red}{red} line highlight the expected coverage while the \textcolor{green}{green} line the real coverage.

```{r filter_muts, include = FALSE}
normal_mutations_filtered <- seq_res_normal %>%
  filter(normal_sample.VAF < 0.8 & normal_sample.VAF >0.2)
```

```{r general_sequencing_normal, echo=FALSE,out.width = '80%',fig.align='center',fig.height=2}
dp <- seq_res_normal %>%
  seq_to_long() %>%
  ggplot(aes(x=DP))+
  ggplot2::geom_histogram()+
  geom_vline(aes(xintercept = mean(DP), color = "Real Mean")) +
    geom_vline(aes(xintercept = 30, color = "Expected Mean")) +
    my_ggplot_theme() +
    scale_color_manual(name = "Statistics", values = c("Real Mean" = "forestgreen", "Expected Mean" = "red"))
dp
```

## Tumour samples

<!-- All of the plots will have a fixed size, and the final dim of the figure is given by -->
<!-- a parameters. -->

```{r calulculate_fig_dims1, echo=TRUE, include=FALSE}
width_single_plot <- 3
height_single_plot <- 2
number_of_samples

### if the plot extends in width 
fix_n_samples_per_width <- 3
fig_width <- 3*3
fig_height <- ceiling(number_of_samples/3)*2
```

Tumour sample were sequenced with a coverage of `r params$sequencing$coverage`x. The plot shows the depth of sequencing of simulated mutations in the tumour samples. The \textcolor{red}{red} line highlight the expected coverage while the \textcolor{green}{green} line the real coverage.

```{r general_sequencing_tumour, echo=FALSE,out.width = '80%',fig.align='center',fig.height=fig_height,fig.width=fig_width}
dp <- seq_res_tumour %>%
  seq_to_long() %>%
  ggplot(aes(x=DP))+
  ggplot2::geom_histogram()+
  geom_vline(aes(xintercept = mean(DP), color = "Real Mean")) +
    geom_vline(aes(xintercept = params$sequencing$coverage, color = "Expected Mean")) +
    my_ggplot_theme() +
    scale_color_manual(name = "Statistics", values = c("Real Mean" = "forestgreen", "Expected Mean" = "red"))+
  facet_wrap(~sample_name)
dp
```

# Copy number profile

This section show copy number profiles of the sequenced tumour samples, which results by the rate of passenger CNAs and the
driver CNA events added during the setup of th `MutationEngine`.

## Genome wide plot of Depth Ratio

In Copy Number Alteration (CNA) analysis, the depth ratio is a crucial metric used to infer copy number variations (CNVs) from sequencing data. It represents the relative read depth of a genomic region compared to a reference, helping to determine duplications, deletions, or normal copy states. Driver events are highlighted.

```{r filter_germline,echo=FALSE, include=FALSE}
## keep only germline mutations for genome-wide plots
g_seq <- left_join(normal_mutations_filtered,seq_res_tumour,
                   by =c("chr","chr_pos", "ref", "alt","causes","classes")) %>%
      as.data.frame()
```


```{r calulculate_fig_dims2, echo=TRUE, include=FALSE}
width_single_plot <- 8
height_single_plot <- 2

### if the plot extends in width 
fix_n_samples_per_height <- 3
fig_height <- number_of_samples*height_single_plot
fig_width <- ceiling(number_of_samples/3)*8
```

```{r depth_ratio, echo=FALSE,out.width = '100%',fig.align='center',fig.height=fig_height,fig.width=fig_width}
dr_plots <- lapply(sample_names, function(s){
      dr <- plot_DR_n(seq_res = g_seq,sample = s)
})
wrap_plots(dr_plots,ncol=ceiling(number_of_samples/3),nrow = fix_n_samples_per_height)
```

## Genome wide plot of B Allele Frequency

B Allele Frequency (BAF) is a key metric used in Copy Number Alteration (CNA) analysis to assess allelic imbalances. It represents the proportion of sequencing reads supporting the B allele (one of the two inherited alleles) at heterozygous single nucleotide polymorphisms (SNPs).

```{r baf, echo=FALSE,out.width = '100%',fig.align='center',fig.height=fig_height,fig.width=fig_width}

baf_plots <- lapply(sample_names, function(s){
      baf <- plot_BAF_n(seq_res = g_seq,sample = s)
})
wrap_plots(baf_plots,ncol=ceiling(number_of_samples/3),nrow = fix_n_samples_per_height)
```

## Genome wide plot of Variant Allele Frequency

```{r vaf, echo=FALSE,out.width = '100%',fig.align='center',fig.height=fig_height,fig.width=fig_width}
s_seq <- seq_res_tumour %>%
  filter(classes!="germinal")
vaf_plots <- lapply(sample_names, function(s){
      vaf <- plot_VAF_n(seq_res = s_seq,sample = s)
})
wrap_plots(vaf_plots,ncol=ceiling(number_of_samples/3),nrow = fix_n_samples_per_height)
```



## Segmented Data

```{r segmented_data, echo=FALSE,out.width = '100%',fig.align='center',fig.height=fig_height,fig.width=fig_width}
segmented_plot <- plot_clone_segments(files_cna)
segmented_plot
```

# Somatic mutations

## Clonal e Subclonal Variant Allele Frequency Spectra

## Multivariate VAF across samples

