---
title: "Plot and annotate a forest"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Plot and annotate a forest}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  markdown:
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r message=FALSE, warning=FALSE}
require(rRACES)
```

# Generate a simulation

## Initialize a population

Generate simulation with one population `A`, growth rate and death rates 0.08 and 0.01, respectively.
Get a sample after 200 cell divisions.

```{r}
sim = new(Simulation)
sim$add_mutant(name = "A", growth_rates = 0.08, death_rates = 0.01)
sim$place_cell("A", 500, 500)
sim$run_up_to_time(200)
sim$sample_cells("MySample", c(500, 500), c(510, 510))
```

```{r}
sim
```


## Add mutations

Generate a demo mutation engine. Set the passenger rate of SNVs to 1e-9 and set a driver SNV and a driver CNA.
```{r}
m_engine = build_mutation_engine(setup_code = "demo")

m_engine$add_mutant(mutant_name = "A",
                    passenger_rates = c(SNV = 1e-9),
                    drivers = list(SNV("22", 10510210, "C"), 
                                   CNA(type = "A", chr = "22",
                                       chr_pos = 10303470, len = 200000)))

m_engine
```

Add exposures SBS13 and SBS1 for the first 50 cell divisions, and SBS17b and SBS3 for the last 100.
```{r}
m_engine$add_exposure(coefficients = c(SBS13 = 0.2, SBS1 = 0.8))
m_engine$add_exposure(time=50, coefficients = c(SBS17b = 0.2, SBS3 = 0.8))
m_engine
```

## Sample a forest

Sample a forest and place mutations caused by drivers and mutational signatures as well as 1000 pre-neoplastic mutations.
```{r}
forest = sim$get_samples_forest()
forest

forest_muts = m_engine$place_mutations(forest, 1000)
forest_muts
```

# Plot the forest and annotate events

First, plot the forest. Note that you can only plot a forest with no mutations.
```{r message=FALSE, warning=FALSE}
tree_plot = plot_forest(forest)
tree_plot
```

## Annotate the forest with sampling time, MRCAs, signatures and drivers

```{r message=FALSE, warning=FALSE}
annotate_forest(tree_plot, forest_muts, samples = T, MRCAs = F, exposures = F, drivers=F)
annotate_forest(tree_plot, forest_muts, samples = T, MRCAs = T, exposures = F, drivers=F)
annotate_forest(tree_plot, forest_muts, samples = T, MRCAs = T, exposures = T, drivers=F)
annotate_forest(tree_plot, forest_muts, samples = T, MRCAs = T, exposures = T, facet_signatures = F, drivers=F)
annotate_forest(tree_plot, forest_muts, samples = T, MRCAs = T, exposures = T, drivers=T, add_driver_label = F)
annotate_forest(tree_plot, forest_muts, samples = T, MRCAs = T, exposures = T, drivers=T, add_driver_label = T)
```

